\name{}: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-04-10 paper/main.tex
.

The most common scheme for drug--target interaction prediction is the
split over drug-target pairs \cite{Survey2018} where it may happen
that most drugs and targets that are including in the model's
validation and testing phase have also been included in the training
phase (as part of other drug--target pairs). This scheme is prone to a
number of biases. If the number of interactions for a drug or protein
are imbalanced, i.e., some drugs or proteins have many more
interactions than others, these will be seen more often during
training and they will likely also have more interactions in the
testing and validation sets. Preferentially predicting

The second most prevalent arrangement is the split
over drugs, while only close to none is aiming on a protein split.
However, the first and second splitting scheme are exposed to the
first dataset bias and are hence more likely vulnerable to
transductive inference by just predicting recently seen structures,
rather than implementing inductive inference and generalizing over the
drug representations. Second, these two strategies are more
susceptible to the second bias, as only in these cases the model may
overfit on the number of existing interactions for a single protein,
while in the third scheme the number of interactions of the test
proteins is entirely unknown during training process.

>>>>

The most common scheme for drug--target interaction prediction is the
split over drug-target pairs \cite{Survey2018} where it may happen
that most drugs and targets that are including in the model's
validation and testing phase have also been included in the training
phase (as part of other drug--target pairs). This scheme is prone to a
number of biases. If the number of interactions for a drug or protein
are imbalanced, i.e., some drugs or proteins have many more
interactions than others, these will be seen more often during
training and they will likely also have more interactions in the
testing and validation sets; because some entities have more
interactions, i.e., they are more likely to interact, any model that
preferentially predicts these as interaction partners will improve its
predictive performance. While this accurately captures the
distribution, predicting based on biases in the number of interaction
partners is not desirable when applying the model to novel
entities. We have demonstrated that even a ``na\"ive'' classifier that
makes predictions only by exploiting the imbalanced number of
interaction partners can achieve performance close to state of the art
methods (when measuring Macro AUC). When training a machine learning
model on such imbalanced data, it will eventually overfit to this
imbalance.  Splitting by entity (protein or drug) can reduce the
impact of this spurious correlation but not reduce it entirely,
because similar entities will still exhibit similar interaction
patterns. In our experiments, we observed the impact of splitting
training and evaluation sets by protein as a decrease in overall
performance (Macro AUC), providing some evidence that models trained
using this splitting scheme are less sensitive to overfitting to this
type of bias.

The way in which training and evaluation data is generated is related
to how the model is evaluated. An evaluation based on Macro AUC
evaluated the application scenario where a set of drugs and proteins
are given, and out of all possible pairs, the more likely interactions
need to be identified. However, this does not correspond to most
scenarios for drug repurposing where a drug that targets a {\em
  specific} protein (e.g., a protein involved in a disease) needs to
be identified. We use an evaluation measure based on micro-averaged
per protein (Micro AUC$_p$) to evaluate this application.

Finally, a potential source of differences in evaluation is how
negatives are treated during evaluation (and training).
Random sampling of unknown...

% The second most prevalent arrangement is the split
% over drugs, while only close to none is aiming on a protein split.
% However, the first and second splitting scheme are exposed to the
% first dataset bias and are hence more likely vulnerable to
% transductive inference by just predicting recently seen structures,
% rather than implementing inductive inference and generalizing over the
% drug representations. Second, these two strategies are more
% susceptible to the second bias, as only in these cases the model may
% overfit on the number of existing interactions for a single protein,
% while in the third scheme the number of interactions of the test
% proteins is entirely unknown during training process.
